---
title: "SNP filtering conditions"
output: html_notebook
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/2024-Digitaria/scripts/")
```

### Distribution of mean depth calculated from all the 2-allelic SNPs

Reanalyze the conditions with the sample set after removal of presumably
misidentified samples (DG_a and DG_c). Consider only two bcf files, each of
which contains D. ciliaris and all (D. radicosa and D. ciliaris) samples,
respectively.

Filtering conditions:
 - Discard SNPs with more than 1 missing samples
 - Retain SNPs which has 2 alleles at maximum


```{sh}
export PATH=${HOME}/local/bin:${PATH}

which bcftools
which vcftools
bcftools --version
vcftools --version

cd ~/2024-Digitaria/030-vcf

# D. ciliaris
bcftools view --with-header Dci2.bcf |
vcftools --vcf - --recode --stdout --max-alleles 2 |
vcftools --vcf - --site-mean-depth --stdout --max-missing-count 1 \
  > Dci2.site-mean-depth.tsv

# D. radicosa & D. ciliaris
bcftools view --with-header DG2.bcf |
vcftools --vcf - --recode --stdout --max-alleles 2 |
vcftools --vcf - --site-mean-depth --stdout --max-missing-count 1 \
  > DG2.site-mean-depth.tsv

```

```{r}
library(tidyverse)
depth_Dci <- read.table("../030-vcf/Dci2.site-mean-depth.tsv", header = TRUE, sep = "\t")
depth_Dci <- mutate(depth_Dci, TYPE = "Dci", .before = everything())
depth_DG  <- read.table("../030-vcf/DG2.site-mean-depth.tsv",  header = TRUE, sep = "\t")
depth_DG  <- mutate(depth_DG, TYPE = "DG", .before = everything())

depth <- bind_rows(depth_Dci, depth_DG)
depth <- depth |> 
  mutate(TYPE = factor(TYPE), CHROM = factor(CHROM)) |> 
  group_by(TYPE)
summarise(depth, n())

depth |> 
  ggplot(aes(x = MEAN_DEPTH)) +
  facet_grid(cols = vars(TYPE)) +
  geom_histogram() +
  theme_bw()
```

```{r}
depth |> 
  filter(MEAN_DEPTH < 250) |> 
  ggplot(aes(x = MEAN_DEPTH)) +
  facet_grid(cols = vars(TYPE)) +
  geom_histogram() +
  theme_bw()
```

```{r}
depth |> 
  filter(MEAN_DEPTH < 100) |> 
  ggplot(aes(x = MEAN_DEPTH)) +
  facet_grid(cols = vars(TYPE)) +
  geom_histogram() +
  theme_bw()
```

```{r}
depth |> 
  filter(MEAN_DEPTH <= 50) |> 
  ggplot(aes(x = MEAN_DEPTH)) +
  facet_grid(cols = vars(TYPE)) +
  geom_histogram(binwidth = 1) +
  theme_bw()
```

 -> Based on the observation of these histograms, filtering of SNPs by depth:
    use SNPs with AD in range 5 - 50.

###